---
title: "P8105 HW3"
output: github_document
date: "2024-10-09"
---

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(rnoaa)
library(ggridges)
library(dplyr)
library(ggplot2)
```

```{r}
library(p8105.datasets)
data("ny_noaa")
```

## Problem 1 

```{r}
ny_noaa = 
  (ny_noaa %>%
mutate(
    date = as.Date(date),
    year = year(date),
    month = month(date),
    day = day(date),  
    tmax = as.numeric(tmax),
    tmin = as.numeric(tmin),
    prcp = as.numeric(prcp),
    snow = as.numeric(snow)))
```

In the ny_noaa dataset, we observe weather patterns (precipation, snowfall, snow depth) from numerous stations in New York. The variables identified are: (station) id, date, tmax (max temp in F), tmin (min temp in F), prcp (precipation), and snow (fall). The most commonly observed snowfall values are 0 or NA indicating that there is a large majority of stations that do not have snowfall. 

```{r}
```{r}
snowfall_values <- ny_noaa %>%
  group_by(snow) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

print(snowfall_values)
```


## Problem 2 

